### 2022-12-21 3-reports/characterization_report.Rmd
overall summaries in report

---
title: "Synpotic Soil Characterization"
output: 
  github_document:
  html_preview: true

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

---

## OVERALL SUMMARIES

```{r analyte-completion}
tar_read(analysis_completion_matrix) %>% knitr::kable()
```

```{r overall-pca}
tar_read(gg_pca_all)
```

```{r correlations-all}
tar_read(gg_correlations)
```

---

**Gravimetric moisture**

```{r moisture}
tar_read(gg_moisture)
```

**Loss on ignition (% organic matter)**

```{r loi}
tar_read(gg_loi)
```

**pH**

```{r pH}
tar_read(gg_pH)
```

**specific conductance**

```{r sp_cond}
tar_read(gg_sp_cond)
```

**total CNS**
```{r tctnts}
tar_read(gg_tctnts)
```

**Water extractable organic C**

```{r weoc}
tar_read(gg_weoc)
```

**Extractable inorganic N**

Inorganic N was extracted using 2M KCl

```{r din}
tar_read(gg_din)
```

**ICP results: base cations**

Extracted using 1M NH4Cl

```{r icp-base cations, fig.height=12, fig.width=7}
tar_read(gg_icp)$gg_base_cations
```

**ICP results: other cations**

Extracted using 1M NH4Cl

```{r icp-other cations, fig.height=9, fig.width=7}
tar_read(gg_icp)$gg_other_cations
```

**ICP results: others**

Extracted using 1M NH4Cl

```{r icp-others, fig.height=4, fig.width=7}
tar_read(gg_icp)$gg_others
```

**HCl-extractable iron - ferrozine method**

Extracted using 0.5 M HCl

```{r iron-ferrozine, fig.height=5, fig.width=7}
tar_read(gg_ferrozine)
```

**Mehlich-3 extractable P**

Extracted using Mehlich-3 solution, analyzed with molybdate method

```{r phosphorus-mehlich, fig.height=5, fig.width=7}
tar_read(gg_mehlich)
```

**FT-ICR-MS**

```{r, icr-vk, fig.height=16, fig.width=16}
tar_read(gg_icr_vankrevelen)
```

```{r, icr-unique, fig.height=16, fig.width=16}
tar_read(gg_icr_unique)
```

```{r, icr-pca, fig.height=5, fig.width=5}
tar_read(gg_icr_pca)$biplot_all
```

```{r, icr-pca-regions, fig.height=5, fig.width=10}
tar_read(gg_icr_pca)$biplot_regions
```

```{r, ions, fig.height=16, fig.width=16}
tar_read(gg_ions)
```

---

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>