### 2023-01-06 3-reports/characterization_report.Rmd
update graphs

---
title: "Synpotic Soil Characterization"
output: 
  github_document:
  html_preview: true

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

---

## OVERALL SUMMARIES

### Analysis completion matrix

<details>
  <summary>click to open</summary>

Number of reps available per analysis. Most were 8-9 reps, but some (e.g. WLE B-horizons) had fewer reps.

```{r analyte-completion}
tar_read(analysis_completion_matrix) %>% knitr::kable()
```

</details>

### Overall trends - PCAs and Correlations

<details>
  <summary>click to open</summary>

**PCAs**

```{r overall-pca, fig.height=6, fig.width=6}
tar_read(gg_pca_all)$gg_pca_overall
```

```{r overall-pca-regions, fig.height=6, fig.width=12}
tar_read(gg_pca_all)$gg_pca_regions
```

**CORRELATIONS**

```{r correlations-all, fig.height=8, fig.width=8}
tar_read(gg_correlations)$corr_all
```

```{r correlations-regions, fig.height=8, fig.width=18}
tar_read(gg_correlations)$corr_regions
```

</details>

---

## EACH ANALYSIS

### by transect

<details>
  <summary>click to open</summary>

```{r by_transect, fig.height=4, fig.width=10}
tar_read(gg_by_transect)
```

</details>

### by site - OA horizons

<details>
  <summary>click to open</summary>

```{r by_site_oa, fig.height=4, fig.width=10}
tar_read(gg_by_site_oa)
```

</details>

### by site - OAB horizons

<details>
  <summary>click to open</summary>

```{r by_site_oab, fig.height=4, fig.width=10}
tar_read(gg_by_site_oab)
```

</details>

---

## FT-ICR-MS

<details>
  <summary>click to open</summary>

```{r, icr-vk-domains, fig.height=5, fig.width=5}
tar_read(gg_icr_vankrevelen)$vk_domains
```

```{r, icr-vk-wle, fig.height=5, fig.width=10}
tar_read(gg_icr_vankrevelen)$vk_wle

```

```{r, icr-vk-cb, fig.height=10, fig.width=10}
tar_read(gg_icr_vankrevelen)$vk_cb

```

```{r, icr-unique-wle, fig.height=10, fig.width=4}
tar_read(gg_icr_unique)$vk_unique_wle
```

```{r, icr-unique-cb, fig.height=10, fig.width=7}
tar_read(gg_icr_unique)$vk_unique_cb
```

```{r, icr-pca, fig.height=5, fig.width=5}
tar_read(gg_icr_pca)$biplot_all
```

```{r, icr-pca-regions, fig.height=5, fig.width=10}
tar_read(gg_icr_pca)$biplot_regions
```

</details>

---

## Session Info

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>