### 2024-03-15 3-reports/playing_with_pcas.Rmd
cleaning

---
title: "Playing with PCAs"
output: 
  github_document:
  html_preview: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
source("2-code/0-packages.R")
library(ggbiplot)

pal_transect = rev(soilpalettes::soil_palette("redox2", 4))
theme_set(theme_kp())
```

```{r}
targets::tar_load_everything()
```

```{r}
make_data_subset = function(data_combined){

 # data_combined_subset = 
    data_combined %>% 
    filter(!transect %in% "wte") %>% 
    #filter(!grepl("Bromide|Fluoride|Nitrate|Calcium|Magnesium|Potassium|Sodium|Phosphate|Ammonia", name, ignore.case = TRUE)) %>% 
    force()
}
make_data_wide = function(data_combined_subset, sample_key){

  #data_combined_wide = 
    data_combined_subset %>% 
    dplyr::select(sample_label, analysis, name, value) %>% 
    separate(name, sep = "_", into = "variable", remove = F) %>% 
    #mutate(name = paste0(variable, " (", analysis, ")")) %>% 
    dplyr::select(sample_label, variable, value) %>% 
    pivot_wider(names_from = "variable") %>% 
    left_join(sample_key) %>% 
    filter(!grepl("016", sample_label)) %>% # this one sample is very weird
    force()

}

```

```{r, eval=TRUE}

data_combined_subset = make_data_subset(data_combined) # %>% filter(!analysis %in% "LOI")
data_combined_wide_NO_DIC = make_data_wide(data_combined_subset %>% 
                                             filter(!grepl("dic", name, ignore.case = TRUE)),
                                           sample_key)
```

```{r}
## Set df for PCA

data_combined_wide = make_data_wide(data_combined_subset, sample_key)
data_wide_PCA = 
  data_combined_wide %>% 
  dplyr::select(-c("dic", "gwc", "percentOM")) %>% 
  rename(CEC = cec,
         NH4N = nh4n,
         NO3N = no3n,
         P = mehlichp,
         WEOC = npoc)

```

```{r}
fit_pca_function = function(dat){

  dat %>% 
    drop_na()

  num = 
    dat %>%       
    dplyr::select(where(is.numeric)) %>%
    dplyr::mutate(row = row_number()) %>% 
    drop_na()

  num_row_numbers = num %>% dplyr::select(row)

  grp = 
    dat %>% 
    dplyr::select(where(is.character)) %>% 
    dplyr::mutate(row = row_number()) %>% 
    right_join(num_row_numbers)

  num = num %>% dplyr::select(-row)
  pca_int = prcomp(num, scale. = T)

  list(num = num,
       grp = grp,
       pca_int = pca_int)
}

```

---

## OVERALL

```{r}
## PCA input files ----
pca_overall = fit_pca_function(data_wide_PCA) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
```

```{r}
## PCA plots overall ----
#gg_pca_overall = 
  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(#shape = pca_overall$grp$transect,
               color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall$grp$region,
               color = pca_overall$grp$transect))+ 
 # scale_color_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

```

### regions separated - colored by transect

```{r, fig.width=10, fig.height=6}
gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

#  library(patchwork)
#  gg_pca_regions = gg_pca_wle + gg_pca_cb
gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
gg_pca_regions

```

### regions separated - colored by site

```{r, fig.width=10, fig.height=6}
gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$site), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$transect,
                 color = groups))+ 
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$site), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$transect,
                 color = groups))+ 
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

#  library(patchwork)
#  gg_pca_regions = gg_pca_wle + gg_pca_cb
gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
gg_pca_regions

```

---

## BY TRANSECT

```{r}
# BY TRANSECT ----

pca_overall_upland = fit_pca_function(data_wide_PCA %>% filter(transect == "upland"))
pca_overall_transition = fit_pca_function(data_wide_PCA %>% filter(transect == "transition"))
pca_overall_wetland = fit_pca_function(data_wide_PCA %>% filter(transect == "wc"))

gg_overall_upland = 
  ggbiplot(pca_overall_upland$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_upland$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_upland$grp$site,
                 color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  labs(shape="",
       title = "PCA upland, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_overall_transition = 
  ggbiplot(pca_overall_transition$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_transition$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_transition$grp$site,
                 color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  labs(shape="",
       title = "PCA transition, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_overall_wetland = 
  ggbiplot(pca_overall_wetland$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wetland$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wetland$grp$site,
                 color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  labs(shape="",
       title = "PCA wetland, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_overall_upland
gg_overall_transition
gg_overall_wetland

#

```

---

## BY TRANSECT + REGION

```{r, fig.width=10, fig.height=6}
# BY TRANSECT + REGION ----

pca_upland_wle = fit_pca_function(data_wide_PCA %>% filter(transect == "upland" & region == "WLE"))
pca_upland_cb = fit_pca_function(data_wide_PCA %>% filter(transect == "upland" & region == "CB"))

gg_upland_wle = 
  ggbiplot(pca_upland_wle$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_upland_wle$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(shape="",
       title = "PCA upland, WLE",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_upland_cb = 
  ggbiplot(pca_upland_cb$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_upland_cb$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(shape="",
       title = "PCA upland, CB",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

cowplot::plot_grid(gg_upland_wle , gg_upland_cb)

```

```{r, fig.width=10, fig.height=6}

pca_transition_wle = fit_pca_function(data_wide_PCA %>% filter(transect == "transition" & region == "WLE"))
pca_transition_cb = fit_pca_function(data_wide_PCA %>% filter(transect == "transition" & region == "CB"))

gg_transition_wle = 
  ggbiplot(pca_transition_wle$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_transition_wle$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(shape="",
       title = "PCA transition, WLE",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_transition_cb = 
  ggbiplot(pca_transition_cb$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_transition_cb$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(
       title = "PCA transition, CB",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

cowplot::plot_grid(gg_transition_wle , gg_transition_cb)

```

```{r, fig.width=10, fig.height=6}

pca_wetland_wle = fit_pca_function(data_wide_PCA %>% filter(transect == "wc" & region == "WLE"))
pca_wetland_cb = fit_pca_function(data_wide_PCA %>% filter(transect == "wc" & region == "CB"))

gg_wetland_wle = 
  ggbiplot(pca_wetland_wle$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_wetland_wle$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(
       title = "PCA wetland, WLE",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_wetland_cb = 
  ggbiplot(pca_wetland_cb$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_wetland_cb$grp$site), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(
                 color = groups))+ 
  labs(
       title = "PCA wetland, CB",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

cowplot::plot_grid(gg_wetland_wle , gg_wetland_cb)

```

---

## Plotting PC1 (drivers of variability)

```{r}
get_pc_values = function(dat){

  pca = fit_pca_function(dat)

  x = pca$pca_int$rotation
  x2 = 
    x %>% 
    as.data.frame() %>% 
    dplyr::select(PC1) %>% 
    arrange(desc(PC1)) %>% 
    rownames_to_column("variable") %>% 
    mutate(variable = fct_reorder(variable, (PC1)))

  x2
}
```

```{r, fig.height=8, fig.width=6}

pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "WLE")) %>% mutate(region = "WLE")
pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "CB")) %>% mutate(region = "CB")

pc_all = 
  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
  mutate(variable = fct_reorder(variable, (PC1)))

pc_all %>% 
  mutate(variable = fct_reorder(variable, (PC1))) %>% 
  filter(region != "combined") %>% 
  ggplot(aes(x = PC1, y = variable, color = region))+
  geom_point(size = 5)+
  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
  xlim(-0.4, 0.4)+
  geom_vline(xintercept = 0)+
  labs(y = "")+
  theme_kp()

pc_all %>% 
  mutate(variable = fct_reorder(variable, (PC1))) %>% 
  filter(region != "combined") %>% 
  ggplot(aes(x = abs(PC1), y = variable, color = region))+
  geom_point(size = 5)+
  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
 # xlim(0, 0.4)+
  #geom_vline(xintercept = 0)+
  labs(y = "",
       x = "|PC1|")+
  theme_kp()

```

PC1 loadings by transect
```{r}

pc_transect = 
  data_wide_PCA %>% 
  group_by(transect) %>% 
  do(get_pc_values(.)) %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc")))

pc_transect %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable))+
  geom_point(size = 5)+
  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
 # xlim(0, 0.4)+
  #geom_vline(xintercept = 0)+
  labs(y = "",
       x = "|PC1|")+
  facet_grid(. ~ transect)+
  theme_kp()

```

PC1 loadings by region+transect
```{r}

pc_region_transect = 
  data_wide_PCA %>% 
  group_by(region, transect) %>% 
  do(get_pc_values(.)) %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc")))

pc_region_transect %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable, color = region))+
  geom_point(size = 5)+
  scale_colour_manual(values = c("#81C0A7", "#1A421D"))+
 # xlim(0, 0.4)+
  #geom_vline(xintercept = 0)+
  labs(y = "",
       x = "|PC1|")+
  facet_grid(. ~ transect)+
  theme_kp()

```

---

---

---

## Session Info

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>