### 2024-08-20 3-reports/characterization_report.Rmd
skipping psa and xrd plots for now. #5

---
title: "Synpotic Soil Characterization"
output: 
  github_document:
  html_preview: true

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

---

## OVERALL SUMMARIES

### Analysis completion matrix

<details>
  <summary>click to open</summary>

Number of reps available per analysis. Most were 8-9 reps, but some (e.g. WLE B-horizons) had fewer reps.

```{r analyte-completion}
tar_read(analysis_completion_matrix) %>% knitr::kable()
```

</details>

### Overall trends - PCAs and Correlations

<details>
  <summary>click to open</summary>

**PCAs**

```{r overall-pca, fig.height=6, fig.width=6}
tar_read(gg_pca_all)$gg_pca_overall
```

```{r overall-pca-regions, fig.height=6, fig.width=12}
tar_read(gg_pca_all)$gg_pca_regions
```

**CORRELATIONS**

```{r correlations-all, fig.height=8, fig.width=8}
tar_read(gg_correlations)$corr_all
```

```{r correlations-regions, fig.height=8, fig.width=18}
tar_read(gg_correlations)$corr_regions
```

</details>

---

## EACH ANALYSIS

### figures for manuscript

<details>
  <summary>click to open</summary>

```{r ms_mehlich, fig.height=5, fig.width=6}
tar_read(gg_by_transect_colorsites)$gg_wle_mehlich
```

```{r ms_wle_icp, fig.height=9, fig.width=9}
tar_read(gg_by_transect_colorsites)$gg_wle_icp
```

```{r ms_cb_icp, fig.height=9, fig.width=9}
tar_read(gg_by_transect_colorsites)$gg_cb_icp
```

```{r ms_wle_anions, fig.height=5, fig.width=9}
tar_read(gg_by_transect_colorsites)$gg_wle_anions
```

```{r ms_cb_anions, fig.height=5, fig.width=9}
tar_read(gg_by_transect_colorsites)$gg_cb_anions
```

```{r ms_cb_fe, fig.height=5, fig.width=6}
tar_read(gg_by_transect_colorsites)$gg_cb_fe
```

```{r ms_cb_cond, fig.height=5, fig.width=6}
tar_read(gg_by_transect_colorsites)$gg_cb_spcond
```

</details>

### by transect

<details>
  <summary>click to open</summary>

```{r by_transect, fig.height=4, fig.width=10}
tar_read(gg_by_transect)
```

</details>

### by site - OA horizons

<details>
  <summary>click to open</summary>

```{r by_site_oa, fig.height=4, fig.width=10}
tar_read(gg_by_site_oa)
```

</details>

### by site - OAB horizons

<details>
  <summary>click to open</summary>

```{r by_site_oab, fig.height=4, fig.width=10}
tar_read(gg_by_site_oab)
```

</details>

---

---

## XRD
```{r xrd, fig.height=4, fig.width=15}
#tar_read(gg_xrd)
```

---

## Soil Physics

### WRC

```{r wrc, fig.height=7, fig.width=8}
tar_read(gg_wrc)

```

### Texture
```{r texture}
#tar_read(gg_texture)
```

---

## Session Info

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>