### 2024-10-23 3-reports/manuscript_figures.Rmd
Fe/P plotting

---
title: "Synpotic Soil Characterization"
output: 
  github_document:
  html_preview: true

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

---

```{r}
source("2-code/0-packages.R")
library(ggbiplot)

pal_transect = rev(soilpalettes::soil_palette("redox2", 4))
theme_set(theme_kp())
```

```{r targets}
targets::tar_load(c(data_combined, sample_key))
```

```{r data_processing}
make_data_subset = function(data_combined){

 # data_combined_subset = 
    data_combined %>% 
    filter(!transect %in% "wte") %>% 
    force()
}
make_data_wide = function(data_combined_subset, sample_key){

  #data_combined_wide = 
    data_combined_subset %>% 
    dplyr::select(sample_label, analysis, name, value) %>% 
    separate(name, sep = "_", into = "variable", remove = F) %>% 
    dplyr::select(sample_label, variable, value) %>% 
    pivot_wider(names_from = "variable") %>% 
    left_join(sample_key) %>% 
    filter(!grepl("016", sample_label)) %>% # this one sample is very weird
    force()

}

data_combined_subset = make_data_subset(data_combined)

data_combined_wide = make_data_wide(data_combined_subset, sample_key) %>% 
  dplyr::select(-gwc) %>% 
    rename(CEC = cec,
         NH4N = nh4n,
         NO3N = no3n,
         P = mehlichp,
         WEOC = npoc,
         DIC = dic,
         SOM = percentOM)

data_wide_PCA = 
  data_combined_wide %>% 
  dplyr::select(-c("DIC", "SOM"))

print("df created: `data_combined_wide`, `data_wide_PCA`")
```

---

## 0. Correlations

```{r gg_corr, fig.height=8, fig.width=8}

fit_correlations_function = function(dat, TITLE){
  num = 
    dat %>%       
    dplyr::select(where(is.numeric)) %>%
    drop_na()

  num_clean = 
    num %>% 
    rownames_to_column("row") %>% 
    pivot_longer(-row) %>% 
    separate(name, sep = "_", into = c("name")) %>% 
    pivot_wider() %>% 
    dplyr::select(-row)

  m = cor(num_clean)
  m = round(m, 1)

  p.mat <- ggcorrplot::cor_pmat(num_clean)
  ggcorrplot::ggcorrplot(m, type = "lower",
                         hc.order = TRUE,
                         p.mat = p.mat,
                         outline.color = "black",
                         lab = TRUE, 
                         insig = "blank",
                         #colors = c("#E46726", "white", "#6D9EC1"),
  ) +
    scale_fill_gradient2(low = "orangered2", mid = "white", high = "turquoise4", limit = c(-1, 1), midpoint = 0)+
    theme(legend.position = "none")

}

corr_all = fit_correlations_function(dat = data_wide_PCA)
#corr_wle = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "WLE"), TITLE = "WLE" )
#corr_cb = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "CB"), TITLE = "CB")

corr_all
```

Highly correlated variables include:

- (+) TC & WEOC, TN
- (+) TS & Chloride, Sulfate, Na, CEC
- (+) CEC & TS, Na, K, Mg, SpConductance
- (+) Na & SpConductance, TS, K, Mg
- (+) Chloride & Sulfate, TS, CEC, SpConductance, Na
- (-) pH & WEOC, Al
- (-) Ca and Al

---

# VERSION 1: ALL ANALYTES

<details>
  <summary> Click to Open </summary>

## 1. PCAs

```{r pca_function}
fit_pca_function = function(dat){

  dat %>% 
    drop_na()

  num = 
    dat %>%       
    dplyr::select(where(is.numeric)) %>%
    dplyr::mutate(row = row_number()) %>% 
    drop_na()

  num_row_numbers = num %>% dplyr::select(row)

  grp = 
    dat %>% 
    dplyr::select(where(is.character)) %>% 
    dplyr::mutate(row = row_number()) %>% 
    right_join(num_row_numbers)

  num = num %>% dplyr::select(-row)
  pca_int = prcomp(num, scale. = T)

  list(num = num,
       grp = grp,
       pca_int = pca_int)
}
```

---

### Overall PCAs

```{r pca_input}
pca_overall = fit_pca_function(data_wide_PCA) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
```

```{r pca_gg, fig.width=10, fig.height=6}
## PCA plots overall
#gg_pca_overall = 
ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_overall$grp$region), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_overall$grp$region,
                 fill = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E"))+
  scale_fill_manual(values = c("#16879C", "#BB281E"))+
  scale_shape_manual(values = c(23, 21))+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  guides(fill = "none")+
  NULL

```

```{r pca_gg_regions, fig.width=10, fig.height=6}

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-6,6)+
  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-5.5,7.5)+
  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb)
gg_pca_regions

```

### Drivers and loadings

```{r pc_loadings}
get_pc_values = function(dat){

  pca = fit_pca_function(dat)

  x = pca$pca_int$rotation
  x2 = 
    x %>% 
    as.data.frame() %>% 
    dplyr::select(PC1) %>% 
    arrange(desc(PC1)) %>% 
    rownames_to_column("variable") %>% 
    mutate(variable = fct_reorder(variable, (PC1)))

  x2
}

pc_overall = get_pc_values(data_wide_PCA) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA %>% filter(region == "WLE")) %>% mutate(region = "WLE")
pc_cb = get_pc_values(data_wide_PCA %>% filter(region == "CB")) %>% mutate(region = "CB")

```

```{r pc_gg, fig.height=8, fig.width=6}

pc_all = 
  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
  mutate(variable = fct_reorder(variable, (PC1)))

overall = 
  pc_all %>% 
  filter(region == "combined") %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+

  xlim(-0.4, 0.4)+
  geom_vline(xintercept = 0)+
  labs(y = "")+
  theme_kp()

non_absolute = 
  pc_all %>% 
  mutate(variable = fct_reorder(variable, (PC1))) %>% 
  filter(region != "combined") %>% 
  ggplot(aes(x = PC1, y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+

  xlim(-0.4, 0.4)+
  geom_vline(xintercept = 0)+
  labs(y = "")+
  theme_kp()

pc_all %>% 
  filter(region != "combined") %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+
  labs(y = "",
       x = "|PC1|")+
  theme_kp()

```

### Clustering

```{r cluster_calc, include=FALSE}

#library(cluster)
library(factoextra)

df = data_wide_PCA %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA %>% drop_na() %>% dplyr::select(!where(is.numeric))

fviz_nbclust(df, FUN = hcut, method = "wss")
fviz_nbclust(df, FUN = hcut, method = "silhouette")
# three clusters

d <- dist(df, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
plot(hc1, cex = 0.6, hang = -1)
hc5 <- hclust(d, method = "ward.D2" )
sub_grp <- cutree(hc5, k = 3)
table(sub_grp)
plot(hc5, cex = 0.6)
rect.hclust(hc5, k = 3, border = 2:5)

df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
count(df_cl,cluster)

```

```{r cluster_gg}
df_cl %>%
  ggplot(aes(y = cluster, x = site))+
  geom_point(aes(color = transect, shape = transect),
             size = 4, stroke = 1,
             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
                                             dodge.width = 0.4))+
  scale_color_manual(values = pal_transect)

```

### PCA with clusters

```{r cluster_pca, fig.width=10, fig.height=6}
library(ggConvexHull)

pca_cluster = fit_pca_function(df_cl)

ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
                   groups = as.character(pca_cluster$grp$cluster), 
                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_convexhull(aes(group = pca_cluster$grp$cluster,
                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_cluster$grp$region,
                 fill = groups))+ 
  scale_shape_manual(values = c(23, 21))+
  scale_color_manual(values = pnw_palette("Sailboat", 3))+
  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
  labs(fill="cluster",
       color = "cluster",
       shape = "region",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

```

---

## 1b. PERMANOVA

```{r}
library(vegan)

adonis2((data_wide_PCA %>% dplyr::select(where(is.numeric)) %>% drop_na()) ~ 
         (region + transect + horizon + site)^2, 
       data = data_wide_PCA %>% drop_na)

```

## 2. Each analyte

### Normalized values - v1

```{r scaled_gg1, fig.height = 8, fig.width = 7}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

data_scaling_max_min = 
  data_long %>% 
  filter(transect != "transition") %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  group_by(region, site, horizon, name) %>% 
  dplyr::summarise(max = max(value, na.rm = T),
                   min = min(value, na.rm = T))

data_scaled2 = 
  data_long %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  left_join(data_scaling_max_min) %>% 
  mutate(scaled = (value - min) / (max - min)) %>% 
  dplyr::select(region, site, transect, horizon, name, scaled) %>%
  pivot_wider(names_from = "transect", values_from = "scaled") %>% 
  mutate(correction = case_when(upland < wetland ~ 1,
                                upland > wetland ~ -1)) %>% 
  pivot_longer(cols = c(upland, wetland, transition), names_to = "transect", values_to = "scaled") %>% 
  mutate(scaled = scaled * correction,
         scaled = case_when(transect == "upland" ~ 0,
                            transect == "wetland" ~ 1,
                            TRUE ~ scaled),
         scaled2 = case_when(scaled > 4 ~ 4,
                             scaled < -1 ~ -1,
                             TRUE ~ scaled),
         label = case_when(scaled > 4 | scaled < -1 ~ round(scaled,1))) %>% 
  # order analytes
  mutate(name = factor(name, levels = c("TC", "TN", "TS", "SOM",
                                        "pH", "spConductance",
                                        "WEOC", "DIC",
                                        "Fe", "P",
                                        "Ca", "Mg", "Na", "K", "Al", "CEC",
                                        "NH4N", "NO3N", "Cl", "SO4")),
         site = factor(site, levels = c("CRC", "PTR", "OWC",
                                        "GCW", "MSM", "GWI")))

data_scaled2 %>%
  ggplot(aes(x = scaled2, y = name))+
  geom_segment(aes(x = 0, xend = scaled2, yend = name), 
               color = "black", linetype = "dashed", linewidth = 0.2)+
  geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
  geom_point(data = data_scaled2 %>% filter(transect != "transition"),
             color = "black",
             size = 1)+
  geom_point(data = data_scaled2 %>% filter(transect == "transition"), 
             size = 4, aes(color = site), stroke = 1, shape = 21, fill = "white")+
  annotate("text", label = "U", x = 0, y = 21)+
  annotate("text", label = "W", x = 1, y = 21)+
  annotate("text", label = "", x = 1, y = 21.5)+
  xlim(-1, 5)+
  scale_y_discrete(limits = rev)+
  scale_shape_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                     values = c(21, 22, 24, 21, 22, 24))+
  labs(x = "",
       y = "",
       shape = "transition soils at")+
  facet_wrap(~region)+
  theme_kp()+
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "right",
        legend.background = element_blank()
  )

```

### Normalized values - v2

```{r scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

data_scaled_NEW = 
  data_long_all %>% 
  group_by(region, site, horizon, transect, name) %>% 
  dplyr::summarise(value_scaled = median(value)) %>% 
  group_by(region, site, horizon, name) %>% 
  dplyr::mutate(value_scaled = scales::rescale(value_scaled)) %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland")))

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

```

```{r scaled_gg2_split, fig.height = 16, fig.width = 18}
library(ggh4x) # for nested facets
# https://teunbrand.github.io/ggh4x/reference/facet_nested.html

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 8,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

### GWI only

```{r scaled_gg2_gwi, fig.height = 8, fig.width = 8}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name + region)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

```

</details>

# VERSION 2: DROPPING CORRELATED ANALYTES

### Overall PCAs

```{r drop}
# drop correlated analytes

data_wide_PCA2 = data_wide_PCA %>% 
  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance))
```

```{r v2-pca_input}
pca_overall = fit_pca_function(data_wide_PCA2) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "CB"))
```

```{r v2-pca_gg, fig.width=10, fig.height=6}
## PCA plots overall
#gg_pca_overall = 
ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
         groups = as.character(pca_overall$grp$region), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_overall$grp$region,
                 fill = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E"))+
  scale_fill_manual(values = c("#16879C", "#BB281E"))+
  scale_shape_manual(values = c(23, 21))+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  guides(fill = "none")+
  NULL

```

```{r v2-pca_gg_regions, fig.width=12, fig.height=7}

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-3, 5)+
  ylim(-2.5, 3.5)+
  labs(shape="",
       title = "PCA: WLE")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
  xlim(-5.5, 3)+
  ylim(-3, 3.5)+
  labs(shape="",
       title = "PCA: CB")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_regions = cowplot::plot_grid(gg_pca_wle , gg_pca_cb, nrow = 1)
gg_pca_regions

```

### Drivers and loadings

```{r v2-pc_loadings}
get_pc_values = function(dat){

  pca = fit_pca_function(dat)

  x = pca$pca_int$rotation
  x2 = 
    x %>% 
    as.data.frame() %>% 
    dplyr::select(PC1) %>% 
    arrange(desc(PC1)) %>% 
    rownames_to_column("variable") %>% 
    mutate(variable = fct_reorder(variable, (PC1)))

  x2
}

pc_overall = get_pc_values(data_wide_PCA2) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA2 %>% filter(region == "WLE")) %>% mutate(region = "WLE")
pc_cb = get_pc_values(data_wide_PCA2 %>% filter(region == "CB")) %>% mutate(region = "CB")

```

```{r v2-pc_gg, fig.height=8, fig.width=6}

pc_all = 
  bind_rows(pc_overall, pc_wle, pc_cb) %>% 
  mutate(variable = fct_reorder(variable, (PC1)))

overall = 
  pc_all %>% 
  filter(region == "combined") %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+

  xlim(-0.4, 0.4)+
  geom_vline(xintercept = 0)+
  labs(y = "")+
  theme_kp()

non_absolute = 
  pc_all %>% 
  mutate(variable = fct_reorder(variable, (PC1))) %>% 
  filter(region != "combined") %>% 
  ggplot(aes(x = PC1, y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+

  xlim(-0.4, 0.4)+
  geom_vline(xintercept = 0)+
  labs(y = "")+
  theme_kp()

pc_all %>% 
  filter(region != "combined") %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+
  labs(y = "",
       x = "|PC1|")+
  theme_kp()

```

### Clustering

```{r v2-cluster_calc, include=FALSE}

#library(cluster)
library(factoextra)

df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))

fviz_nbclust(df, FUN = hcut, method = "wss")
fviz_nbclust(df, FUN = hcut, method = "silhouette")
# three clusters

d <- dist(df, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
plot(hc1, cex = 0.6, hang = -1)
hc5 <- hclust(d, method = "ward.D2" )
sub_grp <- cutree(hc5, k = 3)
table(sub_grp)
plot(hc5, cex = 0.6)
rect.hclust(hc5, k = 3, border = 2:5)

df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp) %>% mutate(cluster = as.character(cluster))
count(df_cl,cluster)

```

```{r v2-cluster_gg}
df_cl %>%
  ggplot(aes(y = cluster, x = site))+
  geom_point(aes(color = transect, shape = transect),
             size = 4, stroke = 1,
             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
                                             dodge.width = 0.4))+
  scale_color_manual(values = pal_transect)

```

### PCA with clusters

```{r v2-cluster_pca, fig.width=10, fig.height=6}
library(ggConvexHull)

pca_cluster = fit_pca_function(df_cl)

ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1, varname.size = 5,
                   groups = as.character(pca_cluster$grp$cluster), 
                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_convexhull(aes(group = pca_cluster$grp$cluster,
                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_cluster$grp$region,
                 fill = groups))+ 
  scale_shape_manual(values = c(23, 21))+
  scale_color_manual(values = pnw_palette("Sailboat", 3))+
  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
  labs(fill="cluster",
       color = "cluster",
       shape = "region",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

```

---

## 1b. PERMANOVA

```{r}
library(vegan)

adonis2((data_wide_PCA2 %>% dplyr::select(where(is.numeric)) %>% drop_na()) ~ 
         (region + transect + horizon + site)^2, 
       data = data_wide_PCA2 %>% drop_na)

```

## 2. Each analyte

### Normalized values - v1

```{r v2-scaled_gg1, fig.height = 8, fig.width = 7, eval=FALSE}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

data_scaling_max_min = 
  data_long %>% 
  filter(transect != "transition") %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  group_by(region, site, horizon, name) %>% 
  dplyr::summarise(max = max(value, na.rm = T),
                   min = min(value, na.rm = T))

data_scaled2 = 
  data_long %>% 
  group_by(region, site, transect, horizon, name) %>% 
  dplyr::summarise(value = mean(value)) %>% 
  left_join(data_scaling_max_min) %>% 
  mutate(scaled = (value - min) / (max - min)) %>% 
  dplyr::select(region, site, transect, horizon, name, scaled) %>%
  pivot_wider(names_from = "transect", values_from = "scaled") %>% 
  mutate(correction = case_when(upland < wetland ~ 1,
                                upland > wetland ~ -1)) %>% 
  pivot_longer(cols = c(upland, wetland, transition), names_to = "transect", values_to = "scaled") %>% 
  mutate(scaled = scaled * correction,
         scaled = case_when(transect == "upland" ~ 0,
                            transect == "wetland" ~ 1,
                            TRUE ~ scaled),
         scaled2 = case_when(scaled > 4 ~ 4,
                             scaled < -1 ~ -1,
                             TRUE ~ scaled),
         label = case_when(scaled > 4 | scaled < -1 ~ round(scaled,1))) %>% 
  # order analytes
  mutate(name = factor(name, levels = c("TC", "TN", "TS", "SOM",
                                        "pH", "spConductance",
                                        "WEOC", "DIC",
                                        "Fe", "P",
                                        "Ca", "Mg", "Na", "K", "Al", "CEC",
                                        "NH4N", "NO3N", "Cl", "SO4")),
         site = factor(site, levels = c("CRC", "PTR", "OWC",
                                        "GCW", "MSM", "GWI")))

data_scaled2 %>%
  ggplot(aes(x = scaled2, y = name))+
  geom_segment(aes(x = 0, xend = scaled2, yend = name), 
               color = "black", linetype = "dashed", linewidth = 0.2)+
  geom_segment(aes(x = 0, xend = 1, yend = name), color = "black")+
  geom_point(data = data_scaled2 %>% filter(transect != "transition"),
             color = "black",
             size = 1)+
  geom_point(data = data_scaled2 %>% filter(transect == "transition"), 
             size = 4, aes(color = site), stroke = 1, shape = 21, fill = "white")+
  annotate("text", label = "U", x = 0, y = 21)+
  annotate("text", label = "W", x = 1, y = 21)+
  annotate("text", label = "", x = 1, y = 21.5)+
  xlim(-1, 5)+
  scale_y_discrete(limits = rev)+
  scale_shape_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), 
                     values = c(21, 22, 24, 21, 22, 24))+
  labs(x = "",
       y = "",
       shape = "transition soils at")+
  facet_wrap(~region)+
  theme_kp()+
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "right",
        legend.background = element_blank()
  )

```

### Normalized values - v2

```{r v2-scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance, SOM, DIC, Al)) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

data_scaled_NEW = 
  data_long_all %>% 
  group_by(region, site, horizon, transect, name) %>% 
  dplyr::summarise(value_scaled = median(value)) %>% 
  group_by(region, site, horizon, name) %>% 
  dplyr::mutate(value_scaled = scales::rescale(value_scaled)) %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
  mutate(name = factor(name, levels = c("TC", "TS", "WEOC",
                                        "Ca", "Na", "Chloride",
                                        "NH4N", "Fe", "P")))

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name)+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

```

```{r v2-scaled_gg2_split, fig.height = 10, fig.width = 14}
library(ggh4x) # for nested facets
# https://teunbrand.github.io/ggh4x/reference/facet_nested.html

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 6,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

### GWI only

```{r v2-scaled_gg2_gwi, fig.height = 8, fig.width = 8}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name )+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

```

# Specific analytes

```{r analytesx-2, fig.height = 4, fig.width = 8, eval=FALSE}
data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")
```

```{r analytesx-3, fig.height = 10, fig.width = 8}

plot_transect_as_x = function(data, YLAB = "", TITLE = "", SUBTITLE = "", SCALES = "free_x"){

  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
    filter(!is.na(value))

  fit_lme_hsd_transect = function(dat){
    # this function will assign post-hoc HSD letters to the transect
    # use lme4::lmer -> multcomp::glht -> multcomp::cld
    library(multcomp)

    lmer = lme4::lmer(value ~ transect + (1|site), data = dat)
    h = summary(glht(lmer, linfct = mcp(transect = "Tukey")), test = adjusted("holm"))
    x = cld(h, decreasing = TRUE) # decreasing order of HSD letters
    x$mcletters$monospacedLetters %>% # sigh
      as.data.frame() %>% 
      rownames_to_column("transect") %>% 
      rename(groups = ".") %>% 
      mutate(groups = str_remove_all(groups, " "),
             groups = toupper(groups))

  }

  # make summary table for HSD letters ----
  hsd_transect_overall = 
    data %>% 
    group_by(region) %>% 
    do(fit_lme_hsd_transect(.))

  ## then calculate the y-axis position for the labels
  lme_y =
    data %>% 
    #group_by(region) %>% 
    dplyr::summarise(max = max(value, na.rm =T),
                     y = max + max/5) %>% 
    ungroup()

  ## combine them
  lme_for_graphs = 
    hsd_transect_overall %>% 
    cross_join(lme_y %>% dplyr::select(-max))

  #
  # graphs ----

  ##  ggplot(data,
  ##         aes(x = transect,
  ##             y = value
  ##             #shape = horizon,
  ##             #alpha = horizon
  ##         )) +
  ##    # plot points
  ##    geom_point(aes(group = site,
  ##                   color = site),
  ##               size = 2.5, stroke = 1,
  ##               position = position_dodge(width = 0.4))+
  ##    # plot HSD letters
  ##    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
  ##    scale_alpha_manual(values = c(1, 0.3))+
  ##    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
  ##    facet_wrap(~region, 
  ##               labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
  ##    theme_kp()+
  ##    labs(x = "",
  ##        # y = YLAB,
  ##        # title = TITLE,
  ##        # subtitle = SUBTITLE,
  ##         shape = "Site",
  ##         alpha = "Horizon",
  ##         color = "Site")+
  ##    theme(legend.position = "right", legend.box = "vertical")+
  ##    NULL

  ggplot(data,
         aes(x = transect,
             y = value
             #shape = horizon,
             #alpha = horizon
         )) +
    # plot points

    geom_violin(aes(fill = transect), alpha = 0.4)+
    geom_jitter(width = 0.1)+
    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
    labs(x = "",
         y = "Extractable P, μg/g")+
    facet_wrap(~region)+
    scale_fill_manual(values = pal_transect)+
    labs(x = "",
         y = YLAB,
         title = TITLE,
         subtitle = SUBTITLE,
         shape = "Site",
         alpha = "Horizon",
         color = "Site")+
    theme(legend.position = "none")+
    NULL

}

gg_fe = plot_transect_as_x(data_combined_wide %>% rename(value = Fe),
                   YLAB = "Extractable Fe, μg/g")

gg_p = plot_transect_as_x(data_combined_wide %>% rename(value = P),
                   YLAB = "Extractable P, μg/g")

gg_weoc = plot_transect_as_x(data_combined_wide %>% rename(value = WEOC),
                   YLAB = "Water-Extractable Organic C, μg/g")

cowplot::plot_grid(gg_fe, gg_p, ncol = 1)

```

---

## Session Info

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>